DOpElib: Differential Equations\\ and Optimization Environment

### 2013-05-10 Examples/PDE/InstatPDE/Example5/content.tex
1) Minor changes in the formatting of the documentation. 2) Added documentation to PDE/StatPDE/Example9
Changed DOpE to DOpElib in the documentation, finished overhaul of PDEStatPDE-Examples

\textbf{General problem description}

\vspace{0.2cm}

@@

\vspace{0.2cm}

\textbf{Program description}

\vspace{0.2cm}

The new feature of this example is the nonhomogeneous righthand side. In examples \ref{PDE_Stat_Laplace_2D} and \ref{PDE_Stat_Laplace_3D}, we regarded stationary problems with nonhomogeneous righthand sides, but up to now, we never involved the time variable into the nonhomogeneity. To do this, DOpE yields a \texttt{SetTime()} function which has to be applied in the \textit{localpde.h} file as well as at the place where the \texttt{RightHandSideFunction} class is declared (here the \textit{myfunctions.h} file.

>>>>

\subsubsection{General problem description}

@@

\subsubsection{Program description}

The new feature of this example is the nonhomogeneous righthand side. In examples \ref{PDE_Stat_Laplace_2D} and \ref{PDE_Stat_Laplace_3D}, we regarded stationary problems with nonhomogeneous righthand sides, but up to now, we never involved the time variable into the nonhomogeneity. To do this, \texttt{DOpElib} yields a \texttt{SetTime()} function which has to be applied in the \textit{localpde.h} file as well as at the place where the \texttt{RightHandSideFunction} class is declared (here the \textit{myfunctions.h} file.

### 2013-05-10 Examples/PDE/StatPDE/Example1/content.tex
Changed DOpE to DOpElib in the documentation, finished overhaul of PDEStatPDE-Examples

\vspace{0.2cm}
\index{functional evaluation}
Here, we declare all quantities of interest (functionals), e.g. point values, drag, lift, mean values of certain quantities over a subdomain etc. \\
Each of these functionals is declared as a class of its own, but in DOpE all classes are derived from a so-called \texttt{FunctionalInterface} class.\\
In the current example we declare functionals for point values of the velocity and for the flux at the outflow boundary of the channel.\index{functional!point value} \index{functional!boundary flux}\\

@@

The \textit{LocalPDE} is derived from a \texttt{PDEInterface} class. It comprises several functions which build up the cell and boundary equations, matrices and righthand sides. The weak formulation of problem \eqref{eq:strong} with $f=0$ is
\begin{equation} \label{eq:weak}
   \frac{1}{2}(\nabla v, \nabla \varphi)_\Omega + \frac{1}{2}(\nabla v^{T}, \nabla \varphi)_\Omega - (p, \nabla \cdot \varphi)_\Omega + (\nabla \cdot v, \psi)_\Omega - (n\cdot \nabla v^T,\phi)_{\Gamma_{out}} = 0.
\end{equation}
\begin{remark}
Note the additional term on $\Gamma_{out}$, which is a consequence of the use of the symmetric stress tensor together with the free outflow condition.
\end{remark}
This problem is vector valued, i.e. the velocity variable $v$ has two components and the pressure variable $p$ is a scalar. For the implementation, we use a vector valued solution variable with three components, where the distinction between velocity and pressure is done by use of the deal.II \texttt{FEValuesExtractors} class. \index{vector-valued problem}\\
Furthermore, in DOpE we always interpret the problems in the context of a Newton method. Usually, a PDE in its weak formulation is given as
\begin{equation*}
   a(u;\varphi) = f(\varphi).
\end{equation*}
The lefthand side is implemented in the \texttt{CellEquation} function, the righthand side is implemented in the \texttt{CellRightHandSide} function (which is unused in this example, because $f=0$).
\begin{remark}
The weak formulation might contain some terms on faces or (parts of) the boundary. DOpE is able to handle these via \texttt{BoundaryEquation}, \texttt{BoundaryRightHandSide} etc.. To keep things simple, we neglect these terms in this introduction.
\end{remark}
To apply Newton's method, this problem is linearized: on the lefthand side, we have the derivative of the (semilinear) form $a(\cdot;\cdot)$ with respect to the solution variable $u$, and the righthand side is the residual of the weak formulation:
\begin{equation*}
   a_u'(u;u^+,\varphi) = -a(u;\varphi) + f(\varphi).
\end{equation*}
In the \texttt{CellMatrix} function, we implement the following matrix $A$ as representation of the derivative on the lefthand side:
\begin{equation*}
  A = (a_u'(u;\varphi_i,\varphi_j))_{j,i=1}^N
\end{equation*}
with the number $N$ of the degrees of freedom. Similarly, the \texttt{CellEquation} contains the vector
\begin{equation*}
  a = a(u;\varphi_i)_{i=1}^{N},
\end{equation*}
and the \texttt{CellRightHandSide} in the case $f \neq 0$ would contain a vector
\begin{equation*}
  \tilde{f} = (f;\varphi_i)_{i=1}^N.
\end{equation*}
The system of equations which is then actually solved is
\begin{equation*}
  A\tilde{u}^+ = -a + \tilde{f}.
\end{equation*}
Because of the linearity of equation \eqref{eq:weak}, there is almost no difference between the two functions.\index{Newton's method}\\

>>>>

\vspace{0.2cm}
\index{functional evaluation}
Here, we declare all quantities of interest (functionals), e.g. point values, drag, lift, mean values of certain quantities over a subdomain etc. \\
Each of these functionals is declared as a class of its own, but in \texttt{DOpElib} all classes are derived from a so-called \texttt{FunctionalInterface} class.\\
In the current example we declare functionals for point values of the velocity and for the flux at the outflow boundary of the channel.\index{functional!point value} \index{functional!boundary flux}\\

@@

The \textit{LocalPDE} is derived from a \texttt{PDEInterface} class. It comprises several functions which build up the cell and boundary equations, matrices and righthand sides. The weak formulation of problem \eqref{eq:strong} with $f=0$ is
\begin{equation} \label{eq:weak}
   \frac{1}{2}(\nabla v, \nabla \varphi)_\Omega + \frac{1}{2}(\nabla v^{T}, \nabla \varphi)_\Omega - (p, \nabla \cdot \varphi)_\Omega + (\nabla \cdot v, \psi)_\Omega - (n\cdot \nabla v^T,\phi)_{\Gamma_{out}} = 0.
\end{equation}
\begin{remark}
Note the additional term on $\Gamma_{out}$, which is a consequence of the use of the symmetric stress tensor together with the free outflow condition.
\end{remark}
This problem is vector valued, i.e. the velocity variable $v$ has two components and the pressure variable $p$ is a scalar. For the implementation, we use a vector valued solution variable with three components, where the distinction between velocity and pressure is done by use of the deal.II \texttt{FEValuesExtractors} class. \index{vector-valued problem}\\
Furthermore, in \texttt{DOpElib} we always interpret the problems in the context of a Newton method. Usually, a PDE in its weak formulation is given as
\begin{equation*}
   a(u;\varphi) = f(\varphi).
\end{equation*}
The lefthand side is implemented in the \texttt{CellEquation} function, the righthand side is implemented in the \texttt{CellRightHandSide} function (which is unused in this example, because $f=0$).
\begin{remark}
The weak formulation might contain some terms on faces or (parts of) the boundary. DOpE is able to handle these via \texttt{BoundaryEquation}, \texttt{BoundaryRightHandSide} etc.. To keep things simple, we neglect these terms in this introduction.
\end{remark}
To apply Newton's method, this problem is linearized: on the lefthand side, we have the derivative of the (semilinear) form $a(\cdot;\cdot)$ with respect to the solution variable $u$, and the righthand side is the residual of the weak formulation:
\begin{equation*}
   a_u'(u;u^+,\varphi) = -a(u;\varphi) + f(\varphi).
\end{equation*}
In the \texttt{CellMatrix} function, we implement the following matrix $A$ as representation of the derivative on the lefthand side:
\begin{equation*}
  A = (a_u'(u;\varphi_i,\varphi_j))_{j,i=1}^N
\end{equation*}
with the number $N$ of the degrees of freedom. Similarly, the \texttt{CellEquation} contains the vector
\begin{equation*}
  a = a(u;\varphi_i)_{i=1}^{N},
\end{equation*}
and the \texttt{CellRightHandSide} in the case $f \neq 0$ would contain a vector
\begin{equation*}
  \tilde{f} = (f;\varphi_i)_{i=1}^N.
\end{equation*}
The system of equations which is then actually solved is
\begin{equation*}
  A\tilde{u}^+ = -a + \tilde{f}.
\end{equation*}
Because of the linearity of equation \eqref{eq:weak}, there is almost no difference between the two functions.\index{Newton's method}\\
